---
title: "0409 LE 1st try"
author: "Brick"
date: "2024-04-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Now we are trying to apply Laplacian eigen map to the real data. See the first application on iris data. 
```{r message=FALSE, warning=FALSE}
# install.packages("Rdimtools")
```
# 3D scatter plot
See 3D scatter plot of the original iris data(take the first 3 variables.)

```{r}
library(Rdimtools)
library(plotly)
## use iris data
data(iris)
set.seed(100)
subid = sample(1:150,150)
X     = as.matrix(iris[subid,1:4])
lab   = as.factor(iris[subid,5])
df_3d <- as.data.frame(X)
df_3d$lab <- lab

fig_3d <- plot_ly(df_3d, x=~Sepal.Length, y=~Sepal.Width,z=~Petal.Length,color=~lab,size = c(0.8),colors = c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951'))
fig_3d <- fig_3d %>% add_markers()
fig_3d <- fig_3d %>% layout(title = '3D scatter plot of original iris dataset')

fig_3d
```

# paired 2D scatter plots
See 2D scatter plots for all 6 pairs of variables.

```{r}
varnames <- colnames(df_3d)
fig_list <- list()
count <- 1
for(i in 1:3){
  for (j in (i+1):4){
    fig_2d <- plot_ly(df_3d, x=df_3d[,i], y=df_3d[,j],color=~lab,size = c(0.8),colors = c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951'))
    fig_2d <- fig_2d %>% add_markers()
    fig_2d <- fig_2d %>% layout(title = 'scatter plot of original iris dataset.') 
    fig_list[[count]] <- fig_2d
    count <- count+1
    rm(fig_2d)
  }
}

# Combine all figures into one subplot
combined_fig <- subplot(style(fig_list[[1]], showlegend=FALSE),
                        style(fig_list[[2]], showlegend=FALSE),
                        style(fig_list[[3]], showlegend=FALSE),
                        style(fig_list[[4]], showlegend=FALSE),
                        style(fig_list[[5]], showlegend=FALSE),
                        style(fig_list[[6]], showlegend=FALSE),
                        nrows = 3, margin=0.05)

# Display the combined figure
combined_fig # NOTE! x and y axes need to be added to add readability
```


# Laplacian Eigen map
Use "proportion" to construct neighborhood map

```{r}
## try different levels of connectivity
for (i in c(0.01,0.02,0.03,0.04,0.05,0.10,0.20,0.50,1)){
  le1 <- do.lapeig(X, type=c("proportion",i), weighted=FALSE)
  df_le1 <- as.data.frame(le1$Y)
  colnames(df_le1) <- c("dim1","dim2")
  df_le1$lab <- lab
  fig_le1 <- plot_ly(df_le1, x=~dim1, y=~dim2,color=~lab,size = c(0.8),colors = c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951'))
  fig_le1 <- fig_le1 %>% add_markers()
  fig_le1 <- fig_le1 %>% layout(title = 'Embedded dimension 1 and 2', plot_bgcolor = "#e5ecf6", xaxis = list(title = 'Embedded dim 1'), 
           yaxis = list(title = 'Embedded dim 2'), legend = list(title=list(text='<b> Species of Iris </b>')))
  print(fig_le1)  
}
```
Use "knn(k nearest neighbor)" to construct neighborhood map.

```{r}
## try different levels of connectivity
for (k in seq(10,150,10)){
  le1 <- do.lapeig(X, type=c("knn",k), weighted=FALSE)
  df_le1 <- as.data.frame(le1$Y)
  colnames(df_le1) <- c("dim1","dim2")
  df_le1$lab <- lab
  fig_le1 <- plot_ly(df_le1, x=~dim1, y=~dim2,color=~lab,size = c(0.8),colors = c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951'))
  fig_le1 <- fig_le1 %>% add_markers()
  fig_le1 <- fig_le1 %>% layout(title = paste('Embedded dimension 1 and 2, k=',k), plot_bgcolor = "#e5ecf6", xaxis = list(title = 'Embedded dim 1'), 
           yaxis = list(title = 'Embedded dim 2'), legend = list(title=list(text='<b> Species of Iris </b>')))
  print(fig_le1)  
}
```
Use "enn(epsilon nearest neighbor)" to construct neighborhood map.

```{r}
## try different levels of connectivity
for (e in seq(1,10,1)){
  le1 <- do.lapeig(X, type=c("enn",e), weighted=FALSE)
  df_le1 <- as.data.frame(le1$Y)
  colnames(df_le1) <- c("dim1","dim2")
  df_le1$lab <- lab
  fig_le1 <- plot_ly(df_le1, x=~dim1, y=~dim2,color=~lab,size = c(0.8),colors = c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951'))
  fig_le1 <- fig_le1 %>% add_markers()
  fig_le1 <- fig_le1 %>% layout(title = paste('Embedded dimension 1 and 2, enn, e=',e), plot_bgcolor = "#e5ecf6", xaxis = list(title = 'Embedded dim 1'), 
           yaxis = list(title = 'Embedded dim 2'), legend = list(title=list(text='<b> Species of Iris </b>')))
  print(fig_le1)  
}
```



# Check the Swiss roll data

```{r}
load(file="my_Swiss.Rda")
sub_swiss <- swiss[sample(1:2000,500),]
fig_3d <- plot_ly(sub_swiss, x=~x, y=~y,z=~z,color=~t,size = c(0.8),colors = c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951'))
fig_3d <- fig_3d %>% add_markers()
fig_3d <- fig_3d %>% layout(title = '3D scatter plot of Swiss Roll Data')

fig_3d
```


```{r}
fig_list <- list()
count <- 1
for(i in 1:2){
  for (j in (i+1):3){
    fig_2d <- plot_ly(sub_swiss, x=swiss[,i], y=swiss[,j],color=swiss[,4],size = c(0.8),colors = c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951'))
    fig_2d <- fig_2d %>% add_markers()
    fig_2d <- fig_2d %>% layout(title = 'scatter plot of original swiss dataset.') 
    fig_list[[count]] <- fig_2d
    count <- count+1
    rm(fig_2d)
  }
}

combined_fig <- subplot(fig_list)

# Display the combined figure
combined_fig # NOTE! x and y axes need to be added to add readability
```

```{r}
## try different levels of connectivity
for(k in seq(50,500,50)){
le1 <- do.lapeig(sub_swiss, type=c("knn",k), weighted=FALSE)
df_le1 <- as.data.frame(le1$Y)
colnames(df_le1) <- c("dim1","dim2")
df_le1$lab <- sub_swiss$t
fig_le1 <- plot_ly(df_le1, x=~dim1, y=~dim2,color=~lab,size = c(0.8),colors = c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951'))
fig_le1 <- fig_le1 %>% add_markers()
fig_le1 <- fig_le1 %>% layout(title = paste('Embedded dimension 1 and 2, k=',k), plot_bgcolor = "#e5ecf6", xaxis = list(title = 'Embedded dim 1'), 
         yaxis = list(title = 'Embedded dim 2'), legend = list(title=list(text='<b> Species of Iris </b>')))
print(fig_le1)
}

```


# Check the Corpus data

```{r}
load(file="my_Swiss.Rda")
sub_swiss <- swiss[sample(1:2000,500),]
fig_3d <- plot_ly(sub_swiss, x=~x, y=~y,z=~z,color=~t,size = c(0.8),colors = c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951'))
fig_3d <- fig_3d %>% add_markers()
fig_3d <- fig_3d %>% layout(title = '3D scatter plot of Swiss Roll Data')

fig_3d
```


```{r}
fig_list <- list()
count <- 1
for(i in 1:2){
  for (j in (i+1):3){
    fig_2d <- plot_ly(sub_swiss, x=swiss[,i], y=swiss[,j],color=swiss[,4],size = c(0.8),colors = c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951'))
    fig_2d <- fig_2d %>% add_markers()
    fig_2d <- fig_2d %>% layout(title = 'scatter plot of original swiss dataset.') 
    fig_list[[count]] <- fig_2d
    count <- count+1
    rm(fig_2d)
  }
}

combined_fig <- subplot(fig_list)

# Display the combined figure
combined_fig # NOTE! x and y axes need to be added to add readability
```

```{r}
## try different levels of connectivity
for(k in seq(50,500,50)){
le1 <- do.lapeig(sub_swiss, type=c("knn",k), weighted=FALSE)
df_le1 <- as.data.frame(le1$Y)
colnames(df_le1) <- c("dim1","dim2")
df_le1$lab <- sub_swiss$t
fig_le1 <- plot_ly(df_le1, x=~dim1, y=~dim2,color=~lab,size = c(0.8),colors = c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951'))
fig_le1 <- fig_le1 %>% add_markers()
fig_le1 <- fig_le1 %>% layout(title = paste('Embedded dimension 1 and 2, k=',k), plot_bgcolor = "#e5ecf6", xaxis = list(title = 'Embedded dim 1'), 
         yaxis = list(title = 'Embedded dim 2'), legend = list(title=list(text='<b> Species of Iris </b>')))
print(fig_le1)
}
```



